---
title: "Homework 1"
author: "Jiaqi Xie"
date: "2026-09-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

data("penguins", package = "palmerpenguins")
```


## Problem 1

The `penguins` dataset contains information on penguins from three species:
Adelie, Chinstrap, and Gentoo. Important variables include species, island,
bill length, bill depth, flipper length, body mass, sex, and year.

The dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables.


The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.


```{r}
penguin_plot =
  ggplot(
    penguins,
    aes(
      x = bill_length_mm,
      y = flipper_length_mm,
      color = species
    )
  ) +
  geom_point()

penguin_plot
```
```{r}
ggsave("penguin_scatterplot.pdf", penguin_plot)
```


## Problem 2

```{r, warning = FALSE}
normal_sample = rnorm(10)

sample_df =
  tibble(
    normal_sample = normal_sample,
    positive = normal_sample > 0,
    character_var = letters[1:10],
    factor_var = factor(
      c(
        "low", "medium", "high", "low", "medium",
        "high", "low", "medium", "high", "low"
      )
    )
  )
```

```
  {r, warning = FALSE}
mean(pull(sample_df, normal_sample))
mean(pull(sample_df, positive))
mean(pull(sample_df, character_var))
mean(pull(sample_df, factor_var))
```

The mean can be calculated for the numeric and logical variables, but not
for the character or factor variables. For the logical variable, R treats
TRUE as 1 and FALSE as 0, so the mean represents the proportion of TRUE
values.

```{r, results = "hide", warning = FALSE}
as.numeric(pull(sample_df, positive))
as.numeric(pull(sample_df, character_var))
as.numeric(pull(sample_df, factor_var))
```

Using `as.numeric` on the logical variable converts TRUE to 1 and FALSE to 0.
Applying `as.numeric` to the character variable produces missing values
because the character strings cannot be interpreted as numbers. Applying
`as.numeric` to the factor variable produces numeric codes associated with
the factor levels rather than meaningful quantitative values. This helps
explain why the numeric and logical variables can be averaged, while the
character and factor variables cannot be averaged directly.






